MSc Data Science Project: Can a Convolutional Neural Network judge a book by its cover?

### 2020-09-25 Tex/MSc Data Science Project.tex
Final commit

\documentclass[12pt]{article}
\usepackage{arydshln}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{color}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{mathrsfs}
\usepackage{mdframed}
\usepackage[margin=2cm]{geometry}
\usepackage[superscript,nomove]{cite}
\numberwithin{equation}{section}
\numberwithin{figure}{section}
\usepackage[makeroom]{cancel}
\usepackage{textcomp}
\usepackage{gensymb}
\DeclareMathOperator\arctanh{arctanh}
\usepackage[colorlinks=true]{hyperref}
\usepackage{cleveref}
\crefdefaultlabelformat{[#2#1#3]}
\begin{document}
\renewcommand\citeform[1]{[#1]}
%
\title{MSc Data Science Project: Can a Convolutional Neural Network judge a book by its cover?}
\author{\\Ryan Hill MSci\\\\
Supervisor: Dr Hubie Chen\\\\
Birkbeck, University Of London\\\\
\texttt{rhill06@mail.bbk.ac.uk}}
\date{\today}
\maketitle
\thispagestyle{empty}
\graphicspath{{images/}}
\begin{abstract}
	In this work we revisit the work of Iwana et al.\cite{KenjiIwana} in an attempt to use convolutional neural networks (CNNs) to predict the genre of a book based entirely on its cover image. We test various pre-processing methods for the images and use a selection of newer, more complex, CNNs to attempt to improve on the results of the previous work. Using an Inception-ResNet-v2 we achieve Top-1 accuracy of 24.7\% and a Top-3 accuracy of 45.4\% which is an improvement over previous results.
\end{abstract}
\begin{figure}[!b]
	\centering
	\includegraphics[scale=0.4]{bbk_logo.jpg}
\end{figure}

@@

The question of predicting a genre based only on a cover image is more difficult than it may first appear; many books at least slightly span genres, and designs are at their heart art which means some personal interpretation will always play a part. Being able to successfully build a model to solve this problem would allow designers to use such a tool to ensure that a design is clear to their target audience, it could allow some data-validation of book related data and could help uncover some underlying design principles in genre classification that is not evident to us currently.

In this work we attempt to train a selection of modern CNN architectures that have been pretrained on the ImageNet dataset to correctly predict the genre of a book from its cover image. The remainder of this work is structured as follows; in the remainder of this section we discuss related works, then in \cref{sec:Theory} we discuss some theoretical concepts that we use within the work for those familiar with Machine Learning but not in CNNs. In \cref{sec:Data_collection_and_pre-processing} we present our methods of the adjustment of the existing datasets and the testing of different methods of pre-processing the images. \Cref{sec:Model_Training} contains the information about how we configured and trained the models, followed by the evaluation of our results and discussion of next steps in \cref{sec:Evaluation_and_Further_Exploration} before summarising out work in \cref{sec:Conclusion}. Please note in this work we will use the terms \emph{class, genre, classification,} and \emph{category} interchangeably to describe the genre that a particular book cover belongs to.

\subsection{Related Works} 
\label{sub:Related Works} 
So far as we can tell at the time of writing there is only one piece of prior work that has been completed in the area of predicting book genres based on their cover images; the work completed by Iwana et al.\cite{KenjiIwana}. In their work they produced 2 datasets, the \emph{Book32} dataset and the smaller, balanced, dataset known as \emph{BookCover30}. These datasets are the best collection of cleaned data that are readily available and we use these within our work, with some small adjustments as discussed in \cref{sub:Image Download and Manual Review}. In their work they trained 2 CNNs, LeNet and AlexNet on the BookCover30 dataset, achieving 13.5\% and 24.7\% Top-1 accuracy respectively; more detailed information of their results is available in \cref{tab:results_acc}.

@@

Our work builds atop theirs by using more modern and complex CNNs to attempt to identify if the more powerful architectures are more capable of producing equivalent or improved results whilst training for significantly epochs than the original paper.
% subsection Related Works end 
% section CNN Theory and Architectures
\section{Theory} 
\label{sec:Theory} 
Whilst in this work we do not detail the full design (beyond that provided in \cref{sec:CNN_Architectures}) and reasons for those designs of the models we will use, it may be beneficial for the reader to understand some concepts of neural networks in particular the components that make up convolutional neural networks. In this section we briefly describe some of these components as well as some of the other methods used within this work in \cref{sub:Transfer_Learning,sub:Activation_Maximisation}. We assume the reader already has familiarity with machine learning concepts such as backpropagation and fully connected neural networks.
\subsection{Components of Convolutional Neural Networks} 
\label{sub:Components_of_Convolutional_Neural_Networks} 
\subsubsection{Activation functions}
Most layers of a neural network have some kind of activation function; these are used to ensure both that there is at least some level of activity, and also to place the values within a specific range (usually networks only take in values between 0 and 1 so they need to be in this range). Some normalisation is also usually applied which is discussed later. For our purposes there are only 3 activation functions that are relevant for this work.

@@

\subsubsection{Convolutional layers}
A convolutional layer  is different to the traditional fully connected layers and is used predominantly in image classifications, forming the fundamental building block of a CNN. A convolutional layer is defined by 3 key parameters, and the additional choice of padding. The 3 parameters that define a basic convolutional layer are the number of filters (analogous to the number of neurons in a dense layer) which is also the number of channels the output of that convolution will have, the size of the filters, and the stride (or step size) of the filter.

@@

\begin{equation}
	m_{i,j,k} = \sum_{\chi = 1}^{C}\sum_{\alpha = 1}^{l}\sum_{\beta = 1}^{l} f_{\alpha,\beta, k} \times x_{(\alpha +(i-1)*s), (\beta + (j-1)*s),\chi}
\end{equation}
where $m_{i,j,k}$ is the component of the feature map $\bm{m}_k$ for output channel $k$ at position $(i, j)$, C is the number of channels in the input, $x$ is the 3 dimensional input tensor (height, width, and channels) input, $f$ is the $k^{th}$ kernel of size $l\times l$, $s$ is the stride. The complexity of the subscripts on $x$ arises from the need to describe the offset provided by the size and stride of the kernel. We apply this equation for the examples given in \cref{fig:conv_process} in the following, noticing the simplifications that $C = s = 1$ so that $k \subseteq \{1\}$, and in this case $l = 3$. Then for the examples given in the image we have
\begin{align}
	m_{1, 1, 1} &= \sum_{\alpha = 1}^{l}\sum_{\beta = 1}^{l} f_{\alpha,\beta, 1} \times x_{(\alpha +(1-1)), (\beta + (1-1)),1}	\\
	m_{1, 1, 1} &= (1\times1 )+ (0 \times 2) + (1 \times 1) + (0 \times 2) + (1 \times 0) + (0 \times 0) + (1 \times 1) + (0 \times 0) + (1 \times 2) = 5 \\
	m_{1, 2, 1} &= \sum_{\alpha = 1}^{l}\sum_{\beta = 1}^{l} f_{\alpha,\beta, 1} \times x_{(\alpha +(1-1)), (\beta + (2-1)),1}	\\
	m_{1, 2, 1} &= (1\times 2)+ (0 \times 1) + (1 \times 1) + (0 \times 1) + (1 \times 0) + (0 \times 1) + (1 \times 0) + (0 \times 2) + (1 \times 1) = 3 \\
	\vdots \nonumber	\\
	m_{3, 3, 1} &= \sum_{\alpha = 1}^{l}\sum_{\beta = 1}^{l} f_{\alpha,\beta, 1} \times x_{(\alpha +(3-1)), (\beta + (3-1)),1}	\\
	m_{3, 3, 1} &= (1\times 2)+ (0 \times 1) + (1 \times 0) + (0 \times 0) + (1 \times 2) + (0 \times 1) + (1 \times 1) + (0 \times 0) + (1 \times 2) = 7
\end{align}
Note that in the image the kernel is displayed second visually but due to the order we wrote the equation is in the kernel elements are on the left of the multiplication.

@@

\subsubsection{Depth-Wise Convolutions}
Depth wise convolutions are similar to normal convolutional layers in all ways except one; instead of applying the filter to each channel of the input and summing the outputs to produce a single output channel we instead keep these outputs separate and produce an output with as many channels as the input per filter.  When followed by a 1x1 convolution (also known as a pointwise convolution) this has the effect of merging the layers via addition and combined these two processes actually have a reduced computational cost when compared to a traditional convolution \cite{Sandler2018}. The results from depth-wise convolutions are identical to to traditional convolutions but due to this improvement in computational cost they are used within many complex models to improve performance.

\subsubsection{Pooling}
Pooling is an operation used within convolutional neural networks to reduce the size of channels/feature maps within the network. Pooling is applied using the same size and stride arguments as convolutions however instead of pointwise multiplication they aggregate across their area. These operations tend to either be maximum or average and are determined either by the user or the CNN architecture. These are applied per-channel and no aggregation across channels is completed afterwards. The dimensions of the output after pooling can be calculated using the same method as before displayed in \cref{eq:feat_map2}. Traditionally (and in our work) pooling is completed with no padding. Global pooling refers to when the size of the filter matches the size of the input and a 1x1xdepth output is returned. For example global average pooling on the output feature map of \cref{fig:conv_process} would be a 1x1 result of 7, global average pooling would give 4.333\dots and max pooling of size 2x2 with a stride of 1 would lead to
$$
\begin{bmatrix}
	6 & 6 \\
	6 & 7 \\
\end{bmatrix}
$$
as where average pooling of the same size and stride would result in
$$
\begin{bmatrix}
	5.25 & 4.25 \\
	5.25 & 4.5 \\
\end{bmatrix}
$$.

\subsubsection{Batch Normalisation}
Batch normalisation was introduced in a 2015 paper by Sergey Ioffee and Christain Szegedy\cite{Ioffe2015} and is a step introduced into neural networks, including CNNs, between the layers and the activation functions. Batch normalisation works by applying re-scaling and re-centering transformations to the output of a layer i.e. transforming the output of that layer for that batch to have a mean of 0 and a standard deviation of 1. This has the effect of greatly reducing training time and increases accuracy by way of reducing the change in the data distribution between layers and reducing the change of this over time via backpropagation.

@@

In transfer learning it is possible to take one of two approaches; the first is to use the existing model weight and biases as a starting point for training and to train the entire model using your training data, the second is to freeze the weights and biases for the feature engineering part of the model i.e. all but the fully connected layers at the end, and just train the classification top of the model. In our work we took the second approach as, given the size of the models we are using, training the full model would have been computationally expensive, and given the relatively small amount of training data would also have, when compared to the number of parameters in the model, been unlikely to have improved accuracy.
% subsection Transfer_Learning end

@@

\begin{equation}
	\label{eq:act_max}
	\Delta x_i = -\eta\frac{\partial \hat{y}_t}{\partial x_i} + \mathcal{R}
\end{equation}
where $\eta$ is the learning rate which in this work was set at 1, $\mathcal{R}$ is any regularisation terms, and $\hat{y}_t$ is the activation value of the target class. Simply put, change the values for the input in a way that increases the activation for that class within some constraints. One regularisation term used in this work is L2norm. It is defined as the root mean squared value of the vector and in this case helps ensure to generate a smooth image, penalising edges.
% subsection Activation_Maximisation end 
% section Theory end 
\section{Data collection and pre-processing} 
\label{sec:Data_collection_and_pre-processing} 
\subsection{Image Download and Manual Review} 
\label{sub:Image Download and Manual Review} 
The 207, 572 images within the complete book32 dataset covering all original 32 classes were downloaded at full resolution using a slightly altered version of the script provided within the Iwana repository \cite{iwanarepo}. The original csv containing all image URLs was split into files of no more than 20,000 records each to allow the downloads to be run over multiple sessions due to the slow data transfer rates and the overhead of reading data between Google Drive and Google Colab meaning that even checking before downloading an individual file was not much faster than downloading and writing the file anyway. Each sub-file was then iterated over using the remainder of the original script, downloading each image into a sub-folder of the class for that image i.e. the genre classification. Once all images were downloaded a sample few were manually checked to verify as best as possible that the data was still correct and the downloads has worked.

@@

The models were training once using the training data as described in \cref{sub:Image Download and Manual Review} and validated against the validation data split out at the same point. No cross-validation was done due to resource limitations.
% subsection Configuration end 
\subsection{Training Performance} 
\label{sub:Training_Performance} 
The models were each trained with a maximum of 200 epochs and early stopping with a patience of 20 due to the time constraints; in the end all 3 models resulted in early stopping by epoch 104 at most. The training was completed over multiple sessions due to timeout constraints of the Google Colab platform and the speed of training even using a GPU instance; during each session whenever the validation accuracy of an epoch outperformed the previous best this model was saved with the epoch identifier to Google Drive. When a new session was started the weights and biases of this snapshot were loaded directly in and training we restarted from the same epoch. This does mean that there may be slightly differences compared to had the model just been trained in a single session, but these effects should be minimal.

@@

\section{Evaluation} 
\label{sec:Evaluation_and_Further_Exploration} 
\subsection{Results} 
\label{sub:Results} 
Full results for all 3 models evaluated on the test set are available in \cref{tab:results_acc}, along with those reported by Iwana for ease of comparison. A visual representation of this for our models only is shown in \cref{fig:test_perf}. As a reminder, the dataset used for our models was slightly altered from the original set, and due to the use of a validation dataset was not trained on the fill set of records; but for a comparison it is suitable. Inception-ResNet-v2 is comparable to the results of AlexNet but is slightly more accurate when Top-3 predictions are taken into account. MobileNetV2 is close to LeNet in performance, with just slightly improvements in many genres. ResNeXt50 shows very odd behaviour, with many classes showing near 0\% accuracy even at Top-3, whilst some classes have a very high accuracy such as Biographies and Memoirs. This is likely an artefact of it overlapping specific classes which along with is discussed further in \cref{sub:Further_Analysis}.

@@

\begin{equation}
	d_{colour} = \sqrt{(R_2 - R_1)^2 + ((G_2 - G_1)^2 +((B_2 - B_1)^2 }
\end{equation}
where R/G/B are the average red, green, and blue values for each predicted and true class respectively. We can see a small trends here, particular in MobileNetV2, where the closer the average colour of the true class the better the recall of that class. This is of course a somewhat self-fulfilling prophecy; the more you correctly predict the class the more the dataset will be the true class, however is highlights that even by just guessing based on closeness to average colour you can make some small improvements. This supports the claims made by Iwana in their work.

@@

Finally, we can review the \emph{class overlap} for each model. We define class overlap between class A and B for a single prediction as the ratio of the probability of prediction for class A and class B (where A has a higher probability) i.e.
\begin{equation}
	CO(A, B) \stackrel{\mathrm{def}}{=} \min\left\{\frac{P(A)}{P(B)}, \frac{P(B)}{P(A)}\right\}
\end{equation}
The per genre pair class overlap is defined as the average class overlap across all predictions. As we can see in \cref{fig:class_overlap}, where only class overlaps of greater than 0.5 are displayed, the ResNeXt50 model sees large amounts of class overlap between many classes, which agrees with what we saw before that certain classes were being \emph{confused} for each other by the model. Comparatively, MobileNetV2 sees only a few class overlaps greater than 0.5 and Inception-ResNet-v2 sees none. When we examine those with higher overlap in MobileNetV2 we can begin to understand the potential true overlap between classes; \emph{Self-Help} has overlap with \emph{Health, Fitness and Dieting,} \emph{Science and Maths} has overlap with \emph{Medical Books}, and \emph{Religion and Spirituality} has overlap with \emph{Christian Books and Bibles}. All of these are logical overlaps and were a human tasked with classifying books based on their cover into these categories it is highly probable that different people would classify them differently.

\begin{figure}
	\centering
	\captionsetup{justification=centering}
	\includegraphics[scale=0.6]{class_pred_overlap.png}
	\caption{Average class overlap by model}
	\label{fig:class_overlap}
\end{figure}
% section Further_Analysis end 
\subsection{Feature Visualisation} 
\label{sub:Feature_Visualisation} 
The goal of using feature visualisation was to attempt to understand what, if any, underlying structure or information the models were using to aid their classification. To do this we use activation maximisation as implemented into the python package \emph{tf-keras-vis}\cite{keras-vis} to do the computational work. To produce the images the final layer activation function was changed from softmax to linear to ensure that class is being visualised as opposed to minimising other classes, and the loss function is simply defined as that linear activation (note here that loss is really a misnomer as we are trying to maximise this value as opposed to minimise it as we traditionally would) as discussed in \cref{sub:Activation_Maximisation}. L2 normalisation was used as the default to attempt to avoid pure static images being produced.

We ran the activation maximisation on the Inception-ResNet-v2 model for the 3 best predicted genres (Test Prep, Comics \& Graphic Novels, and Travel) as we believed these had the best probability of returning useful images. The inputs to the model were then trained for 1000 epochs, at which point issues were encountered with the RAM on the machine. During these steps there was some improvement to the loss function however did not appear to converge and when compared to the progress seen on a pure ImageNet model these saw very little movement and remained negative the entire time (where as the examples provided by the package are always positive after the first few epochs). Even when all regularisation terms were removed, allowing the process to fully maximise the activation without constraints we still did not see an improvement in the loss values.

@@

There is still lots to explore in this space as we discussed in \cref{sub:Discussion} and hopefully this work provides a springboard to enable more complex CNNs to be used for genre identification of book covers in the future. The work itself is unlikely to provide any value for real-world use as the accuracy is too low to be of use for either consumers or even cover designers as a form of QA. Genre prediction from cover images remains a complex topic and given the subjectivity of the classifications is unlikely to ever be fully solved, however the work produced here gets a little closer than we were before. So it seems for now that, if you are a CNN, you should follow the original advice and not try to judge a book by its cover.
% section Conclusion end

@@

>>>>

\documentclass[12pt]{article}
\usepackage{arydshln}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{float}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{color}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{mathrsfs}
\usepackage{mdframed}
\usepackage[margin=2cm]{geometry}
\usepackage[superscript,nomove]{cite}
\numberwithin{equation}{section}
\numberwithin{figure}{section}
\usepackage[makeroom]{cancel}
\usepackage{textcomp}
\usepackage{gensymb}
\DeclareMathOperator\arctanh{arctanh}
\usepackage[colorlinks=true]{hyperref}
\usepackage{cleveref}
\crefdefaultlabelformat{[#2#1#3]}
\begin{document}
\renewcommand\citeform[1]{[#1]}
%
\title{MSc Data Science Project: Can a Convolutional Neural Network judge a book by its cover?}
\author{\\Ryan Hill MSci\\\\
Supervisor: Dr Hubie Chen\\\\
Birkbeck, University Of London\\\\
\texttt{rhill06@mail.bbk.ac.uk}}
\date{\today}
\maketitle
\thispagestyle{empty}
\graphicspath{{images/}}
\begin{abstract}
	In this work we revisit and build upon the work of Iwana et al.\cite{KenjiIwana} in an attempt to use convolutional neural networks (CNNs) to predict the genre of a book based entirely on its cover image. We test various pre-processing methods for the images and use a selection of newer, more complex, CNNs to attempt to improve on the results of the previous work. Using an Inception-ResNet-v2 we achieve Top-1 accuracy of 24.7\% and a Top-3 accuracy of 45.4\% which is an improvement over previous results.
\end{abstract}
\renewcommand{\abstractname}{Acknowledgements}
\begin{abstract}
Thank you to Pete Stewart and Openreach Ltd for sponsoring this degree, and thank you David, Scott, and Craig who made working while studying a little easier and less stressful than it would have been otherwise. Thank you to my friends who have put up with me over the last 2 years. Finally most of all I want to thank my parents, who have always supported me and my learning, even though it has been a long time since they understood any of it. 
\end{abstract}

\begin{figure}[!b]
	\centering
	\includegraphics[scale=0.4]{bbk_logo.jpg}
\end{figure}

@@

The question of predicting a genre based only on a cover image is more difficult than it may first appear; many books at least slightly span genres, and designs are at their heart art which means some personal interpretation will always play a part. Being able to successfully build a model to solve this problem would allow designers to use such a tool to ensure that a design is clear to their target audience, it could allow some data-validation of book related data, and could help uncover some underlying design principles in genre classification that is not evident to us currently.

In this work we attempt to train a selection of modern CNN architectures that have been pretrained on the ImageNet dataset to correctly predict the genre of a book from its cover image. The remainder of this work is structured as follows; in the rest of this section we discuss related works, then in \cref{sec:Theory} we discuss some theoretical concepts that we use within the work for those familiar with Machine Learning but not in CNNs. In \cref{sec:Data_collection_and_pre-processing} we present our methods of the adjustment of the existing datasets and the testing of different methods of pre-processing the images. \Cref{sec:Model_Training} contains the information about how we configured and trained the models, followed by the evaluation of our results and discussion of next steps in \cref{sec:Evaluation_and_Further_Exploration} before summarising our work in \cref{sec:Conclusion}. Please note in this work we will use the terms \emph{class, genre, classification,} and \emph{category} interchangeably to describe the genre that a particular book cover belongs to.

\subsection{Related Works} 
\label{sub:Related Works} 
So far as we can tell at the time of writing there is only one piece of prior work that has been completed in the area of predicting book genres based on their cover images; the work published by Iwana et al.\cite{KenjiIwana}. In their work they produced 2 datasets, the \emph{Book32} dataset and the smaller, balanced, dataset known as \emph{BookCover30}. These datasets are the best collection of cleaned data that are readily available and we use these within our work, with some small adjustments as discussed in \cref{sub:Image Download and Manual Review}. In their work they trained 2 CNNs, LeNet and AlexNet on the BookCover30 dataset, achieving 13.5\% and 24.7\% Top-1 accuracy respectively; more detailed information of their results is available in \cref{tab:results_acc}.

@@

Our work builds atop theirs by using more modern and complex CNNs to attempt to identify if the more powerful architectures are more capable of producing equivalent or improved results whilst training for significantly fewer epochs than the original paper.
% subsection Related Works end 
% section CNN Theory and Architectures
\section{Theory} 
\label{sec:Theory} 
Whilst in this work we do not detail the full design (beyond that provided in \cref{sec:CNN_Architectures}) and reasons for those designs of the models we will use, it may be beneficial for the reader to understand some concepts of neural networks in particular the components that make up convolutional neural networks. In this section we briefly describe some of these components as well as some of the other methods used within this work in \cref{sub:Transfer_Learning,sub:Activation_Maximisation}. We assume the reader already has familiarity with introductory machine learning concepts such as backpropagation and fully connected neural networks.
\subsection{Components of Convolutional Neural Networks} 
\label{sub:Components_of_Convolutional_Neural_Networks} 
\subsubsection{Activation functions}
Most layers of a neural network have some kind of activation function (a function that transforms the output of the neuron before passing it to the next layer); these are used to ensure both that there is at least some level of activity, and also to place the values within a specific range (usually networks only take in values between 0 and 1 so they need to be in this range). Some normalisation is also usually applied which is discussed later. For our purposes there are only 3 activation functions that are relevant for this work.

@@

\subsubsection{Convolutional layers}
A convolutional layer  is different to the traditional fully connected layers and is used predominantly in image classification, forming the fundamental building block of a CNN. A convolutional layer is defined by 3 key parameters, and the additional choice of padding. The 3 parameters that define a basic convolutional layer are the number of filters (analogous to the number of neurons in a dense layer) which is also the number of channels the output of that convolution will have, the size of the filters, and the stride (or step size) of the filter.

@@

\begin{equation}
	m_{i,j,k} = \sum_{\chi = 1}^{C}\sum_{\alpha = 1}^{l}\sum_{\beta = 1}^{l} f_{\alpha,\beta, k} \times x_{(\alpha +(i-1)*s), (\beta + (j-1)*s),\chi}
\end{equation}
where $m_{i,j,k}$ is the component of the feature map $\bm{m}_k$ for output channel $k$ at position $(i, j)$, C is the number of channels in the input, $x$ is the 3 dimensional input tensor (height, width, and channels) input, $f$ is the $k^{th}$ filter of size $l\times l$, and $s$ is the stride. The complexity of the subscripts on $x$ arises from the need to describe the offset provided by the size and stride of the kernel. We apply this equation for the examples given in \cref{fig:conv_process} in the following, noticing the simplifications that $C = s = 1$ so that $k \subseteq \{1\}$, and in this case $l = 3$. Then for the examples given in the image we have
\begin{align}
	m_{1, 1, 1} &= \sum_{\alpha = 1}^{l}\sum_{\beta = 1}^{l} f_{\alpha,\beta, 1} \times x_{(\alpha +(1-1)), (\beta + (1-1)),1}	\\
	m_{1, 1, 1} &= (1\times1 )+ (0 \times 2) + (1 \times 1) + (0 \times 2) + (1 \times 0) + (0 \times 0) + (1 \times 1) + (0 \times 0) + (1 \times 2) = 5 \\
	m_{1, 2, 1} &= \sum_{\alpha = 1}^{l}\sum_{\beta = 1}^{l} f_{\alpha,\beta, 1} \times x_{(\alpha +(1-1)), (\beta + (2-1)),1}	\\
	m_{1, 2, 1} &= (1\times 2)+ (0 \times 1) + (1 \times 1) + (0 \times 1) + (1 \times 0) + (0 \times 1) + (1 \times 0) + (0 \times 2) + (1 \times 1) = 3 \\
	\vdots \nonumber	\\
	m_{3, 3, 1} &= \sum_{\alpha = 1}^{l}\sum_{\beta = 1}^{l} f_{\alpha,\beta, 1} \times x_{(\alpha +(3-1)), (\beta + (3-1)),1}	\\
	m_{3, 3, 1} &= (1\times 2)+ (0 \times 1) + (1 \times 0) + (0 \times 0) + (1 \times 2) + (0 \times 1) + (1 \times 1) + (0 \times 0) + (1 \times 2) = 7
\end{align}
Note that in the image the kernel is displayed second visually but due to the order we wrote the equation is in the kernel elements are on the left of the multiplication.

@@

\subsubsection{Depth-Wise Convolutions}
Depth wise convolutions are similar to normal convolutional layers in all ways except one; instead of applying the filter to each channel of the input and summing the outputs to produce a single output channel we instead keep these outputs separate and produce an output with as many channels as the input per filter.  When followed by a 1x1 convolution (also known as a pointwise convolution) this has the effect of merging the layers via addition and combined these two processes actually have a reduced computational cost when compared to a traditional convolution \cite{Sandler2018}. The results from depth-wise convolutions are identical to traditional convolutions but due to this improvement in computational cost they are used within many complex models to improve performance.

\subsubsection{Pooling}
Pooling is an operation used within convolutional neural networks to reduce the size of channels/feature maps within the network. Pooling is applied using the same size and stride arguments as convolutions however instead of pointwise multiplication they aggregate across their area. These operations tend to either be maximum or average and are determined either by the user or the CNN architecture. These are applied per-channel and no aggregation across channels is completed afterwards. The dimensions of the output after pooling can be calculated using the same method as before displayed in \cref{eq:feat_map2}. Traditionally (and in our work) pooling is completed with no padding. Global pooling refers to when the size of the filter matches the size of the input and a 1x1xdepth output is returned. For example global average pooling on the output feature map of \cref{fig:conv_process} would be a 1x1 result of 7, global average pooling would give 4.333\dots and max pooling of size 2x2 with a stride of 1 would lead to
$$
\begin{bmatrix}
	6 & 6 \\
	6 & 7 \\
\end{bmatrix}
$$
as where average pooling of the same size and stride would result in
$$
\begin{bmatrix}
	5.25 & 4.25 \\
	5.25 & 4.5 \\
\end{bmatrix}
$$

\subsubsection{Batch Normalisation}
Batch normalisation was introduced in a 2015 paper by Sergey Ioffee and Christain Szegedy\cite{Ioffe2015} and is a step introduced into neural networks, including CNNs, between the layers and the activation functions. Batch normalisation works by applying re-scaling and re-centering transformations to the output of a layer i.e. transforming the output of that layer for that batch to have a mean of 0 and a standard deviation of 1. This has the effect of greatly reducing training time and increases accuracy by way of reducing the change in the data distribution between different layers and more importantly reducing the change of this over time via backpropagation.

@@

In transfer learning it is possible to take one of two approaches; the first is to use the existing model weights and biases as a starting point for training and to train the entire model using your training data, the second is to freeze the weights and biases for the feature engineering part of the model i.e. all but the fully connected layers at the end, and just train the classification top of the model. In our work we took the second approach as, given the size of the models we are using, training the full model would have been computationally expensive, and given the relatively small amount of training data would also have, when compared to the number of parameters in the model, been unlikely to have improved accuracy.
% subsection Transfer_Learning end

@@

\begin{equation}
	\label{eq:act_max}
	\Delta x_i = -\eta\frac{\partial \hat{y}_t}{\partial x_i} + \mathcal{R}
\end{equation}
where $\eta$ is the learning rate which in this work was set at 1, $\mathcal{R}$ is any regularisation terms, and $\hat{y}_t$ is the activation value of the target class. Simply put, change the values for the input in a way that increases the activation for that class within some constraints. One regularisation term used in this work is L2norm. It is defined as the root mean squared value of the vector and in this case helps ensure to generate a smooth image, penalising edges.
% subsection Activation_Maximisation end 
% section Theory end 
\section{Data collection and pre-processing} 
\label{sec:Data_collection_and_pre-processing} 
\subsection{Image Download and Manual Review} 
\label{sub:Image Download and Manual Review} 
The 207,572 images within the complete book32 dataset covering all original 32 classes were downloaded at full resolution using a slightly altered version of the script provided within the Iwana repository \cite{iwanarepo}. The original csv containing all image URLs was split into files of no more than 20,000 records each to allow the downloads to be run over multiple sessions due to the slow data transfer rates. The overhead of reading data between Google Drive and Google Colab meant that even checking before downloading an individual file if it already existed was not much faster than downloading and writing the file anyway. Each sub-file was then iterated over using the remainder of the original script, downloading each image into a sub-folder of the class for that image i.e. the genre classification. Once all images were downloaded a sample few were manually checked to verify as best as possible that the data was still correct and the downloads has worked.

@@

The models were trained once using the training data as described in \cref{sub:Image Download and Manual Review} and validated against the validation data split out at the same point. No k-fold cross-validation was done due to resource limitations.
% subsection Configuration end 
\subsection{Training Performance} 
\label{sub:Training_Performance} 
The models were each trained with a maximum of 200 epochs and early stopping with a patience of 20 due to the time constraints; in the end all 3 models resulted in early stopping by epoch 104 at most. The training was completed over multiple sessions due to timeout constraints of the Google Colab platform and the speed of training even using a GPU instance; during each session whenever the validation accuracy of an epoch outperformed the previous best this model was saved with the epoch identifier to Google Drive. When a new session was started the weights and biases of this snapshot were loaded directly in and training we restarted from the same epoch. This does mean that there may be slightly differences compared to had the model just been trained in a single session, but these effects should be minimal.

@@

\section{Evaluation} 
\label{sec:Evaluation_and_Further_Exploration} 
\subsection{Results} 
\label{sub:Results} 
Full results for all 3 models evaluated on the test set are available in \cref{tab:results_acc}, along with those reported by Iwana for ease of comparison. A visual representation of this for our models only is shown in \cref{fig:test_perf}. As a reminder, the dataset used for our models was slightly altered from the original set, and due to the use of a validation dataset was not trained on the full set of records; but for a comparison it is suitable. Inception-ResNet-v2 is comparable to the results of AlexNet but is slightly more accurate when Top-3 predictions are taken into account. MobileNetV2 is close to LeNet in performance, with just slightly improvements in many genres. ResNeXt50 shows very odd behaviour, with many classes showing near 0\% accuracy even at Top-3, whilst some classes have a very high accuracy such as Biographies and Memoirs. This is likely an artefact of it overlapping specific classes which along with is discussed further in \cref{sub:Further_Analysis}.

@@

\begin{equation}
	d_{colour} = \sqrt{(R_2 - R_1)^2 + (G_2 - G_1)^2 +(B_2 - B_1)^2 }
\end{equation}
where R/G/B are the average red, green, and blue values for each predicted and true class respectively. We can see a small trends here, particular in MobileNetV2, where the closer the average colour of the true class the better the recall of that class. This is of course a somewhat self-fulfilling prophecy; the more you correctly predict the class the more the dataset will be the true class, however it highlights that even by just guessing based on closeness to average colour you can make some small improvements. This supports the claims made by Iwana in their work.

@@

Finally, we can review the \emph{class overlap} for each model. We define class overlap between class A and B for a single prediction as the ratio of the probability of prediction for class A and class B (where B has a higher probability) i.e.
\begin{equation}
	CO(A, B) \stackrel{\mathrm{def}}{=} \min\left\{\frac{P(A)}{P(B)}, \frac{P(B)}{P(A)}\right\}
\end{equation}
The per genre pair class overlap is defined as the average class overlap across all predictions. As we can see in \cref{fig:class_overlap}, where only class overlaps of greater than 0.5 are displayed, the ResNeXt50 model sees large amounts of class overlap between many classes, which agrees with what we saw before that certain classes were being \emph{confused} for each other by the model. Comparatively, MobileNetV2 sees only a few class overlaps greater than 0.5 and Inception-ResNet-v2 sees none. When we examine those with higher overlap in MobileNetV2 we can begin to understand the potential true overlap between classes; \emph{Self-Help} has overlap with \emph{Health, Fitness and Dieting,} \emph{Science and Maths} has overlap with \emph{Medical Books}, and \emph{Religion and Spirituality} has overlap with \emph{Christian Books and Bibles}. All of these are logical overlaps and were a human tasked with classifying books based on their cover into these categories it is highly probable that different people would classify them differently.

\begin{figure}
	\centering
	\captionsetup{justification=centering}
	\includegraphics[scale=0.6]{class_pred_overlap.png}
	\caption{Average class overlap by model}
	\label{fig:class_overlap}
\end{figure}
% section Further_Analysis end 
\subsection{Feature Visualisation} 
\label{sub:Feature_Visualisation} 
The goal of using feature visualisation was to attempt to understand what, if any, underlying structure or information the models were using to aid their classification. To do this we use activation maximisation as implemented into the python package \emph{tf-keras-vis}\cite{keras-vis} to do the computational work. To produce the images the final layer activation function was changed from softmax to linear to ensure that target class is being visualised as opposed to just minimising minimising other class activations, and the loss function is simply defined as that linear activation (note here that loss is really a misnomer as we are trying to maximise this value as opposed to minimise it as we traditionally would) as discussed in \cref{sub:Activation_Maximisation}. L2 normalisation was used as the default to attempt to avoid pure static images being produced.

We ran the activation maximisation on the Inception-ResNet-v2 model for the 3 best predicted genres (Test Prep, Comics \& Graphic Novels, and Travel) as we believed these had the best probability of returning useful images. The inputs to the model were then trained for 1000 epochs, at which point issues were encountered with the RAM on the machine. During these steps there was some improvement to the loss function however did not appear to converge. When compared to the progress seen on a pure ImageNet model these saw very little movement and remained negative the entire time (where as the examples provided by the package are always positive after the first few epochs). Even when all regularisation terms were removed, allowing the process to fully maximise the activation without constraints we still did not see an improvement in the loss values.

@@

There is still lots to explore in this space as we discussed in \cref{sub:Discussion} and hopefully this work provides a springboard to enable more complex CNNs to be used for genre identification of book covers in the future. The work itself is unlikely to provide any value for real-world use as the accuracy is too low to be of use for either consumers or even cover designers as a form of QA. Genre prediction from cover images remains a complex topic and given the subjectivity of the classifications is unlikely to ever be fully solved, however the work produced here gets a little closer than we were before. So it seems for now that, if you are a CNN, you should follow the old adage and don't judge a book by its cover.
% section Conclusion end

@@

\begin{table}[H]
